Predicting the Housing Market using \\ Linear Regression

### 2023-11-13 manuscript/Manuscript.tex
added to intro

\begin{document}

@@

% An overview of the topic
% Existing works
% Why are existing methods not sufficient? What are the elements of an attractive solution?
% Contributions

@@

\bigskip
\bigskip
\bigskip
\bigskip
\bigskip

@@

This linear regression model aims to predict the house price with the different characteristics of a house by using the trained data and testing it against the testing data. To see how well the model performed in predicting the house prices, R-squared was used and a visualization was created to help see the results. The figure below represents how the model did with predicting the house prices. As we can see, there is a clear trend in the plot showing the accuracy of the model. One noticeable trend in the plot is that the model's predicted value of the house is undervalued compared to the actual price of the house. This trend with the model can be supported with the R-squared value. The calculated R-squared value for the model was 0.6529. An R-squared value of 0.6529 indicates the proportion of the variance in the dependent variable that can be explained by the independent variables in a multiple linear regression model. In other terms, about 65.29\% of the variability in the observed data can be accounted for by the model.

@@

@@

\section{References}
\label{sec:refer}

>>>>

\doublespacing
\begin{document}

@@

The United States has been through three recessions in the last decade. And during that decade, house prices have reached the very bottom of the market and also reached the very top of the market. The market has tripled since 1992. After a brief dip in 2002-2003, they bounced back and skyrocketed over 20\% from May 2003 to November 2004. \cite{Jacobsen2005} Because of the constant change, new models for predicting the housing market also need to change so that the model can stay up to date with the changes that occur. This is important because it gives people insights into whether it's a good time to buy or sell a property, helps policymakers make informed decisions, and allows investors to strategize their moves. House price predictions can help the developer determine the selling price of a house and can also help the customer arrange the right time to purchase a house. \cite{alfiyantin2017}

Engineers in the past had to face a significant challenge in forecasting and assessing house prices. Fortunately, with technological advances, machine learning algorithms offer a solution, enabling users to predict house prices by training data sets on various algorithms. The market is currently flooded with software and models designed to enhance the accuracy of house price predictions, minimizing errors to negligible levels. These predictions can be based on the analysis of different aspects of a house, such as kitchen size, bedroom count, bathroom count, and overall land area. These aspects of a house can go on forever. Analysts employ various techniques to compute and analyze data,  to ensure that the predictions are aligned closely with previous estimations. \cite{singh2021}

Existing methods for predicting housing prices often rely on traditional statistical models that may struggle to capture the complexity and non-linear relationships present in real estate markets. Additionally, these models may not effectively incorporate the wide range of features that influence property values. An attractive solution would involve leveraging advanced machine learning techniques, such as multiple linear regression, to handle the complexity and non-linearity of the data. This allows the model to learn intricate patterns and relationships that may be challenging for traditional methods. Additionally, a comprehensive solution should consider factors such as location, economic indicators, property characteristics, and even social trends to build a more accurate and robust predictive model. The key is to strike a balance between model complexity and interpretability while ensuring it can effectively handle the diverse and dynamic nature of the real estate market.
% An overview of the topic
% Existing works
% Why are existing methods not sufficient? What are the elements of an attractive solution?
% Contributions
The main contribution of this paper is to corroborate the findings of prior research papers on this topic.

@@

@@

This linear regression model aims to predict the house price with the different characteristics of a house by using the trained data and testing it against the testing data. To see how well the model performed in predicting the house prices, R-squared was used and a visualization was created to help see the results. The figure below represents how the model did with predicting the house prices. As we can see, there is a clear trend in the plot showing the accuracy of the model. One noticeable trend in the plot is that the model's predicted value of the house is undervalued compared to the actual price of the house. This trend with the model can be supported with the R-squared value.

@@

The calculated R-squared value for the model was 0.6529. An R-squared value of 0.6529 indicates the proportion of the variance in the dependent variable that can be explained by the independent variables in a multiple linear regression model. In other terms, about 65.29\% of the variability in the observed data can be accounted for by the model.

@@
